A thermodynamic view of dusty protoplanetary disks: axisymmetric stability

### 2016-10-19 result.tex
fixed some typos, corrected kappa^2 dependence on dust

%When $\tstop=0$ we may construct exact dusty equilibra by specifying
%the dust-to-gas ratio, as described in \S\ref{eqm}.

We present numerical solutions to the linearized
equations to examine the effect of dust on the VSI. 
We consider perfectly coupled dust ($\tstop=0$) so that the 
equilibria defined in \S\ref{eqm} are exact steady states.

For simplicity, we mostly consider constant midplane
dust-to-gas ratios $\epsilon_0$ and its characteristic height
$H_\epsilon$. The latter is specified via the dust layer thickness
$\Hd(r_0)$. (We will allow for radial dependices in $\epsilon_0$ and
$H_\epsilon$ in \S\ref{varHd}.)   
These, together with the perturbation wavenumber $k$, comprises main the
parameters of the linear problem. We choose a power-law 
midplane gas density profile, $\rho_\mathrm{g0}\propto r^q$ with $p =
-1.5$. The power-law index for the 
temperature profile is $q=-1$; and we set the gas disk aspect-ratio
$h_\mathrm{g}=0.05$. These are fiducial values used by 
\citet[][\citetalias{lin15} in this section]{lin15}.

Two boundary conditions are needed for the $\tstop=0$
problem. This is expected since this problem is equivalent to
adiabatic hydrodynamics \citep[e.g.][]{lubow93}. We 
impose solid boundaries so that $\delta v_z(\pm\zmax)=0$.  
We solve the linearized equations as a generalized eigenvalue problem 
using a pseudo-spectral code adapted from \cite{lin15}. Amplitudes 
are expanded in Chebyshev polynomials up to order $N_z=512$. We check
results using Eq. \ref{vsi_check}.

\subsection{Qualitative expectations}\label{vsi_est}
\cite{lin15} found the appropriate way to compare 
destabilizing vertical shear to stabilizing vertical buoyancy is
$r\p_z\Omega/\OmK$ against $N_z^2/\OmK^2$. We see from \S\ref{vertshear}
and \S\ref{vbuoyancy} that $|r\p_z\Omega|\sim q h_\mathrm{g}\OmK$ for a
thin, gas dominated disk; while $N_z^2\sim \epsilon\OmK^2$. Thus we
expect dust-induced buoyancy forces to stabilize the disk against the
VSI where $\epsilon \gtrsim h_\mathrm{g}$.

>>>>

We now present numerical solutions to the linearized equations in a 
vertically stratified disk. We consider perfectly coupled dust 
($\tstop=0$) so that the equilibria defined in \S\ref{eqm} are exact
steady states. In this limit any growing modes must be associated with
the imposed temperature gradient (\S\ref{limits}). We thus 
examine the effect of dust-loading on the VSI, previously studied in 
pure gas disks \citep[][\citetalias{lin15} in this section]{lin15}.

For simplicity, we mostly consider constant midplane
dust-to-gas ratios $\epsilon_0$ and its characteristic height
$H_\epsilon$. The latter is specified via the dust layer thickness
$\Hd(r_0)$. (We will allow for radial dependices in $\epsilon_0$ and
$H_\epsilon$ in \S\ref{varHd}.)   
These, together with the perturbation wavenumber $k$, comprises main the
parameters of the linear problem. We choose a power-law 
midplane gas density profile, $\rho_\mathrm{g0}\propto r^p$ with $p =
-1.5$. The fiducial power-law index for the 
temperature profile is $q=-1$; and we set the gas disk aspect-ratio
$h_\mathrm{g}=0.05$. These values were also used by 
\citetalias{lin15}.

Two boundary conditions are needed for the $\tstop=0$
problem \citep[e.g.][]{lubow93}. We impose solid boundaries so that
$\delta v_z(\pm\zmax)=0$. We solve the linearized equations as a
generalized eigenvalue problem using a pseudo-spectral code adapted
from \citetalias{lin15}. Amplitudes are expanded in Chebyshev
polynomials up to order $N_z=512$. We check results using
Eq. \ref{vsi_check}.

\subsection{Qualitative expectations}\label{vsi_est}
\citetalias{lin15} found the appropriate way to compare 
destabilizing vertical shear to stabilizing vertical buoyancy is
$r\p_z\Omega/\OmK$ against $N_z^2/\OmK^2$. We see from \S\ref{vertshear}
and \S\ref{vbuoyancy} that $|r\p_z\Omega|\sim q h_\mathrm{g}\OmK$ for a
thin, gas dominated disk; while $N_z^2\sim \epsilon\OmK^2$. Thus we
expect dust-induced buoyancy forces to stabilize the disk against the
VSI where $\epsilon \gtrsim h_\mathrm{g}$.
